Experiments with Synthetic Data

### 2022-07-13 dev/notebooks/synthetic.qmd
work on streamlining code for experiments

@@

```{julia}
include("dev/utils.jl")
using CounterfactualExplanations, Flux, Plots, PlotThemes, Random, LaplaceRedux, LinearAlgebra
theme(:wong)
output_path = output_dir("synthetic")
www_path = www_dir("synthetic")
```

@@

```{julia}
Random.seed!(123)
using MLJ
N = 1000
X, ys = make_blobs(N, 2; centers=2, as_table=false, center_box=(-2 => 2), cluster_std=0.1)
ys .= ys.==2
X = X'
xs = Flux.unstack(X,2)
data = zip(xs,ys)
plt = plot()
plot_data!(plt,X',ys)
```

@@

```{julia}
nn = Chain(Dense(2,1))
loss(x, y) = Flux.Losses.logitbinarycrossentropy(nn(x), y) 
```

```{julia}
using Flux.Optimise: update!, ADAM
opt = ADAM()
epochs = 10
function _forward_nn(nn, data, epochs)
    for epoch = 1:epochs
        for d in data
            gs = gradient(Flux.params(nn)) do
            l = loss(d...)
            end
            update!(opt, Flux.params(nn), gs)
        end
    end
    return nn
end
_forward_nn(nn, data, epochs)
```

```{julia}
using CounterfactualExplanations.Models
import CounterfactualExplanations.Models: logits, probs # import functions in order to extend
import Main.Models: retrain

# Step 1)
struct NeuralNetwork <: Models.AbstractFittedModel
    model::Any
end

# Step 2)
logits(M::NeuralNetwork, X::AbstractArray) = M.model(X)
probs(M::NeuralNetwork, X::AbstractArray)= σ.(logits(M, X))

# Step 3)
function retrain(M::NeuralNetwork, data; n_epochs=10, τ=1.0) 
    nn = M.model
    nn = _forward_nn(nn, data, n_epochs)
    M = NeuralNetwork(nn)
    return M
end

M = NeuralNetwork(nn)
```

```{julia}
#| echo: false
# Plot the posterior distribution with a contour plot.
plt_nn = plot_contour(X',ys,M;zoom=0);
savefig(plt_nn, joinpath(www_path, "nn_contour.png"))
```

![](www/nn_contour.png)

@@

```{julia}
λ = 0.1
la = Laplace(nn, λ=λ)
BayesLaplace.fit!(la, data)
```

```{julia}
# Step 1)
struct LaplaceNeuralNetwork <: Models.AbstractFittedModel
    la::BayesLaplace.Laplace
end

# Step 2)
logits(M::LaplaceNeuralNetwork, X::AbstractArray) = M.la.model(X)
probs(M::LaplaceNeuralNetwork, X::AbstractArray)= BayesLaplace.predict(M.la, X)

# Step 3)
function retrain(M::LaplaceNeuralNetwork, data; n_epochs=10, τ=1.0) 
    nn = M.la.model
    nn = _forward_nn(nn, data, n_epochs)
    la = Laplace(nn, λ=λ)
    BayesLaplace.fit!(la, data)
    M = LaplaceNeuralNetwork(la)
    return M
end

Mᴸ = LaplaceNeuralNetwork(la)
```

```{julia}
#| echo: false
# Plot the posterior distribution with a contour plot.
plt = plot_contour(X',ys,Mᴸ;zoom=0)
savefig(plt, joinpath(www_path, "la_contour.png"))
```

![](www/la_contour.png)

@@

```{julia}
# Models:
models = (Bayesian=Mᴸ, Plugin=M)
# Generators:
generators = (Bayesian=GreedyGenerator(), Plugin=GenericGenerator(ϵ=0.01))
```

@@

```{julia}
cb = false
γ = 0.75
mod = Mᴸ
gen = generators[:Plugin]
plt_original = plt_nn = plot_contour(X',ys,mod;zoom=0,colorbar=cb,title="(a)")
```

```{julia}
counterfactual_data = CounterfactualData(X,ys')
μ = 0.10
Random.seed!(123)
candidates = findall(ys.==0)
chosen = rand(candidates, Int(round(μ*length(candidates))))
X′ = copy(X)
y′ = copy(ys)
using CounterfactualExplanations.Counterfactuals: counterfactual, counterfactual_label
for i in chosen
    x = X[:,i]
    outcome = generate_counterfactual(x, 1, counterfactual_data, mod, gen,γ=γ)
    X′[:,i] = counterfactual(outcome)
    y′[i] = counterfactual_label(outcome)
end
plt_single = plot_contour(X′',y′,mod;zoom=0,colorbar=cb,title="(b)")
```

@@

```{julia}
xs′ = Flux.unstack(X′,2)
data = zip(xs′,y′)
mod = retrain(mod, data)
plt_single_retrained = plot_contour(X′',y′,mod;zoom=0,colorbar=cb,title="(c)")
```

@@

```{julia}
i = 2
while i <= 10
    counterfactual_data = CounterfactualData(X′,y′')
    Random.seed!(123)
    candidates = findall(y′.==0)
    chosen = rand(candidates, Int(round(μ*length(candidates))))
    xs′ = Flux.unstack(X′,2)
    data = zip(xs′,y′)
    mod = retrain(mod, data)
    for i in chosen
        x = X′[:,i]
        outcome = generate_counterfactual(x, 1, counterfactual_data, mod, gen; γ=γ)
        X′[:,i] = counterfactual(outcome)
        y′[i] = counterfactual_label(outcome)
    end
    i += 1
end
plt_single_repeat = plot_contour(X′',y′,mod;zoom=0,colorbar=cb,title="(d)")
```

>>>>

```{julia}
using Pkg; Pkg.activate("dev")
```

@@

```{julia}
include("dev/utils.jl")
using AlgorithmicRecourseDynamics
using CounterfactualExplanations, Flux, Plots, PlotThemes, Random, LaplaceRedux, LinearAlgebra
theme(:wong)
output_path = output_dir("synthetic")
www_path = www_dir("synthetic")
```

@@

```{julia}
Random.seed!(1234)
using MLJ
N = 1000
X, ys = make_blobs(N, 2; centers=2, as_table=false, center_box=(-2 => 2), cluster_std=0.1)
ys .= ys.==2
X = X'
xs = Flux.unstack(X,2)
data = zip(xs,ys)
counterfactual_data = CounterfactualData(X,ys')
```

@@

```{julia}
using Flux
model = Chain(Dense(2,1))
M = FluxModel(model)
```

```{julia}
n_epochs = 10
using AlgorithmicRecourseDynamics.Models
M = Models.train(M,counterfactual_data; n_epochs=n_epochs)
```

```{julia}
plt_nn = plot(M,counterfactual_data;zoom=0)
savefig(plt_nn, joinpath(www_path, "nn_contour.png"))
```

![](../artifacts/upload/www/synthetic/nn_contour.png)

@@

```{julia}
λ = 0.1
model_laplace = Laplace(M.model, λ=λ)
LaplaceRedux.fit!(model_laplace, data)
```

```{julia}
M_laplace = Models.LaplaceModel(model_laplace)
```

```{julia}
plt_laplace = plot(M_laplace, counterfactual_data;zoom=0)
savefig(plt_laplace, joinpath(www_path, "la_contour.png"))
```

![](../artifacts/upload/www/synthetic/la_contour.png)

@@

```{julia}
# Models:
models = (Bayesian=M_laplace, Plugin=M)
# Generators:
generators = (
    Bayesian=GreedyGenerator(loss=:logitbinarycrossentropy), 
    Plugin=GenericGenerator(loss=:logitbinarycrossentropy,ϵ=0.01)
)
```

@@

```{julia}
cb = false
γ = 0.75
mod = M_laplace
gen = generators[:Plugin]
plt_original = plot(mod,counterfactual_data;zoom=0,colorbar=false,title="(a)")
```

```{julia}
μ = 0.10
Random.seed!(123)
candidates = findall(ys.==0)
chosen = rand(candidates, Int(round(μ*length(candidates))))
X′ = copy(X)
y′ = copy(ys)
using CounterfactualExplanations.Counterfactuals: counterfactual, counterfactual_label
for i in chosen
    x = X[:,i]
    outcome = generate_counterfactual(x, 1, counterfactual_data, mod, gen,γ=γ)
    X′[:,i] = counterfactual(outcome)
    y′[i] = counterfactual_label(outcome)
end
counterfactual_data′ = CounterfactualData(X′,y′')
plt_single = plot(mod,counterfactual_data′;zoom=0,colorbar=false,title="(b)")
```

@@

```{julia}
mod = Models.train(mod, counterfactual_data′; n_epochs=n_epochs)
plt_single_retrained = plot(mod,counterfactual_data′;zoom=0,colorbar=false,title="(c)")
```

@@

```{julia}
i = 2
while i <= 10
    counterfactual_data′ = CounterfactualData(X′,y′')
    Random.seed!(123)
    candidates = findall(y′.==0)
    chosen = rand(candidates, Int(round(μ*length(candidates))))
    mod = Models.train(mod, counterfactual_data′; n_epochs=n_epochs)
    for i in chosen
        x = X′[:,i]
        outcome = generate_counterfactual(x, 1, counterfactual_data, mod, gen; γ=γ)
        X′[:,i] = counterfactual(outcome)
        y′[i] = counterfactual_label(outcome)
    end
    i += 1
end
plt_single_repeat = plot(mod,counterfactual_data′;zoom=0,colorbar=false,title="(d)")
```
